---
title: "Classwork Activity"
author: "cristian santiago"
output:
  html_document:
    df_print: paged
---

# Instructions (Step by Step)



## Step 1: Create fake data
```{r}
set.seed(2073)

students <- data.frame(
  Name = paste("Student", 1:20),
  StudyHours = round(runif(20, 1, 10), 1),
  Satisfaction = round(rnorm(20, mean = 7, sd = 1.5), 1),
  Major = sample(c("Data Science", "Business Analytics", "Math", "Computer Science"), 20, replace = TRUE)
)

head(students)
```

# Remove the Name column

```{r}

Students_Clean <- students
Students_Clean$Name <- NULL

head(Students_Clean)
```


### Question: Why might someone drop a column like name or ID from a dataset?





### Answer: Names or IDs usually don’t help with analysis because they aren’t numeric or useful for patterns. Removing them helps keep the data cleaner and easier to work with.

```{r}
library(ggplot2)
```

## Step 3: visualize

```{r}
ggplot(Students_Clean, aes(x = Major, y = Satisfaction, fill = Major)) +
  geom_boxplot() +
  labs(
    title = "Student Satisfaction by Major",
    x = "Major",
    y = "Satisfaction (1–10)"
  ) +
  theme_minimal()

```
## Step 4: Explain: Write 2-3 scenteces each 
### Questions:
1. The chart shows that satisfaction is pretty similar across majors, but some have slightly higher or lower medians.

2. This could be because different majors have different workloads or stress levels.

3. The chart is somewhat accurate, but the dataset is small and randomly generated.

4. Because the data is fake and only has 20 students, it may not reflect real-world patterns.

## Step 5: Questions on Git, GitHub and R Markdown 
### Questions: 

1. What does git add do? stages files so they are ready to be saved in a commit.

2. What is the difference between git commit and git push?git commit saves changes locally, while git push sends those changes to GitHub.

3. Why is it useful to use GitHub for your code or data projects? GitHub is useful because it stores your work online and lets you track changes and share projects.

4. How could version control help when collaborating? Version control helps teams avoid overwriting each other’s work and keeps history of changes.

5. What is the difference between a code chunk and regular text in R Markdown?A code chunk runs code, while regular text explains what’s happening.

6. How could R Markdown help you explain your analysis to someone else? R Markdown lets you combine code, output, and explanations in one document.





















































