---
title: "Iris Dataset Statistics Lab"
author: "Cristian Santiago"
output:
  html_document:
    df_print: paged
---

# Libraries
```{r}
library(tidyverse)
```

# Load the iris dataset


```{r}
data(iris)
```

```{r} 
# Changing names of each column
colnames(iris) <- c("SepalLen", "SepalWid", "PetalLen", "PetalWid", "Species")
```

```{r} 
# checking  first few rows
head(iris)
```

## Questions:
 1. How many numeric columns are in the dataset?
 
 Four Columns
 
 2. How many species are included?
 
 Three species
 
 3. What is the range of SepalLen? (min and max)
 
 Min: 4.4 Max:7.9
 

# Summary Statistics
Calculate mean, median, variance, SD for numeric columns

```{r}
sapply(iris[,1:4], mean)
sapply(iris[,1:4], median)
sapply(iris[,1:4], var)
sapply(iris[,1:4], sd)
```

## Questions:
1. Which variable has the largest spread?
PetalLen has the largest spread.

2. Which has the smallest spread?
SepalWid has the smallest spread.

3. How does the mean compare to the median for PetalLen?
The mean and median for PetalLen are pretty close, so the data is fairly balanced.

4. What does a large variance indicate?
A large variance means the data is more spread out.

# Summary by Species

```{r}
aggregate(SepalLen ~ Species, data=iris, FUN=mean)
aggregate(PetalWid ~ Species, data=iris, FUN=var)
```

## Questions:
 1. Which species has the largest SepalLen on average? 
 Virginica has the largest SepalLen on average.
 
 
 2. Which species shows the most variability in PetalWid? 
 Virginica shows the most variability in PetalWid.
 
 
 3. What does this tell you about the characteristics of different species 
 This shows that different species have different sizes and variation, which helps tell them apat.
 

# Visualizations

### Histogram of SepalLen

```{r}
hist(iris$SepalLen,
     main="Sepal Length Distribution",
     xlab="Sepal Length (cm)",
     col="pink",
     border="black")
```

## Questions:
 1. What is the shape of the distribution (symmetric, skewed)?  this distribution shows a light right skew.
 2. Are there any unusually low or high values?
 There are no obvious outliers, just some small values that slightly stand out.

# Boxplot of PetalLen by Species

```{r}
boxplot(PetalLen ~ Species,
        data=iris,
        main="Petal Length by Species",
        ylab="Petal Length (cm)",
        col=c("pink","skyblue","purple"))
```

## Questions:
 1. Which species has the longest petals?
 Virginica has the longest petals.
 
 2. Are there any outliers visible?
 Its not very visible if there are outliers, but the species are very different.
 
 3. How do the distributions differ between species?
 They look very different with setosa having the smallest petals and virginica the largest.

# Correlation Matrix

```{r}
cor_matrix <- cor(iris[,1:4])
cor_matrix
```

## Questions:
 1. Which two variables are most strongly correlated?
 PetalLen and PetalWid are the most strongly correlated.
 
 2. Are there any weak or negative correlations?
 SepalWid has weaker and slightly negative correlations with some variables.
 
 3. How might these correlations help distinguish species?
 These correlations help separate species because petal size is very different across them.

# Interpretation 

Write a short paragraph (3-5 sentences) answering:

- Which variables best distinguish species?
- How could a scientist use these statistics in real life?
- Do the histograms, boxplots, and correlations tell a consistent story?

PetalLen and PetalWid are the best for telling the species apart because they look really different for each group. A scientist could use these measurements to figure out what type of plant they’re looking at or notice patterns. The graphs and correlations all point to the same thing, which is that petal size matters the most.



